
```{r}
library(boot)
library(mosaicData)
```

```{r}
plot(trees$Volume ~ trees$Height, main = "Black Cherry Tree Volume Relationship",
     xlab = "Height", ylab = "Volume", pch = 16, col = "blue")
plot(trees$Volume ~ trees$Girth, main = "Black Cherry Tree Volume Relationship", xlab = "Girth", ylab = "Volume", pch = 16, col = "blue")
```
```{r}
sampleMean <- function(data, indices){
  return(mean(data[indices]))
}
```


```{r}
#the following function, volumeEstimate, will index the given data,
#and run several regressions and report and store their r-squared statistics.
volumeEstimate <- function(data, indices){
  d <- data[indices, ]
  H_Relationship <- lm(d$Volume ~ d$Height, data = d)
  H_Relationship.rsq <- summary(H_Relationship)$r.square
  G_Relationship <- lm(d$Volume ~ d$Girth, data = d)
  G_Relationship.rsq <- summary(G_Relationship)$r.square
  G_H_ratio <- d$Girth / d$Height
  G_H_relationship <- lm(d$Volume ~ G_H_ratio, data = d)
  G_H_relationship.rsq <- summary(G_H_relationship)$r.square
  combined_relationship <- lm(d$Volume ~ d$Height + d$Girth, data = d)
  combined_relationship.rsq <- summary(combined_relationship)$r.squared
  combined_relationship2 <- lm(d$Volume ~ d$Height + d$Girth + G_H_ratio)
  combined_relationship2.rsq <- summary(combined_relationship2)$r.squared
  relationships <- c(H_Relationship.rsq,
                     G_Relationship.rsq,G_H_relationship.rsq,
                     combined_relationship.rsq,combined_relationship2.rsq)
  return(relationships)
}
```

#Now that we have our function in place, let's do a bootstrap!
```{r}
results <- boot(data = trees, statistic = volumeEstimate, R = 5000)
#This will give us a bootstrapped result of our statistic of interest,
#volumeEstimate, given the dataset (trees), over the number of iterations (5000).
print(results)
```
#tn* is the rsquared statistic for each regression. Bias and standard error are also reported.

#Histogram of the of the results of the first regression
```{r}
plot(results, index = 1)
```
#Confidence Intervals
```{r}
confidence_interval_H = boot.ci(results, index = 1, conf = 0.95, type = "bca")
ci_H = confidence_interval_H$bca[, c(4, 5)]
print(ci_H)
```
#Histograms
```{r}
hist(results$t[,1], main = "Coefficient of Determination: Height", xlab = "R-Squared", col = "grey")
hist(results$t[,1], main = 'Coefficient of Determination: Height', xlab = 'RSquared',
col = 'grey', prob = T)
lines(density(results$t[,1]), col = 'blue')
abline(v = ci_H, col = 'red')
```

#We can also index the results to see the results we want...what were the results on the 5,000 iterations for the height only regression?
```{r}
results_firstRegression <- results$t[ , 1]
#This will give us the R-squared for each iteration of the first regression.
```

