#Lab 1 (ISLR): Subset Selection Methods

```{r}
library(ISLR)
fix(Hitters)
names(Hitters)
dim(Hitters)
sum(is.na(Hitters$Salary))
```
```{r}
Hitters <- na.omit(Hitters)
dim(Hitters)
sum(is.na(Hitters))
```
```{r}
library(leaps)
regfit.full <- regsubsets(Salary ~ ., Hitters)
summary(regfit.full)
```
```{r}
regfit.full <- regsubsets(Salary ~ ., data = Hitters, nvmax = 19)
reg.summary <- summary(regfit.full)
names(reg.summary)
```
```{r}
reg.summary$rsq
```
#Plot RSS, adjusted r-squared, C_p, and BIC for all models at once to decide which model to select.
```{r}
par(mfrow = c(2, 2))
plot(reg.summary$rss, xlab = "Number of Varaibles", ylab = "RSS", type = "l")
plot(reg.summary$adjr2, xlab = "Number of Variables", ylab = "ADjusted RSq", type = "l")
```
#Plot the maximum
```{r}
par(mfrow = c(2, 2))
plot(reg.summary$rss, xlab = "Number of Varaibles", ylab = "RSS", type = "l")
plot(reg.summary$adjr2, xlab = "Number of Variables", ylab = "ADjusted RSq", type = "l")
which.max(reg.summary$adjr2)
#The max is 11.
points(11, reg.summary$adjr2[11], col = "red", cex = 2, pch = 20)
```
#We can plot the C_p and BIC statistics in a similiar fashion. We are looking for the minimum in this case for both variables.
```{r}
plot(reg.summary$cp, xlab = "Number of Variables", ylab = "Cp", type = "l")
which.min(reg.summary$cp)
#The minimum is 10...let's graph that.
points(10, reg.summary$cp[10], col = "red", cex = 2, pch = 20)
#Do the same for BIC now
which.min(reg.summary$bic)
#The minimum for BIC is 6.
plot(reg.summary$bic, xlab = "Number of Variables", ylab = "BIC", type = "l")
points(6, reg.summary$bic[6], col = "red", cex = 2, pch = 20)
```
#regsubsets package can also do this for us.
```{r}
plot(regfit.full, scale = "r2")
plot(regfit.full, scale = "adjr2")
plot(regfit.full, scale = "Cp")
plot(regfit.full, scale = "bic")
```
```{r}
coef(regfit.full,6)
```

#Forward/Backward stepwise selection
```{r}
regfit.fwd <- regsubsets(Salary ~ ., data = Hitters, nvmax = 19, method = "forward")
summary(regfit.fwd)
regfit.bwd <- regsubsets(Salary ~ ., data = Hitters, nvmax = 19, method = "backward")
summary(regfit.bwd)
```
```{r}
coef(regfit.full, 7)
coef(regfit.fwd, 7)
coef(regfit.bwd, 7)
```
#Choosing among models using the validaiton set approach with cross-validation
```{r}
set.seed(1)
train <- sample(c(TRUE, FALSE), nrow(Hitters), rep = TRUE)
test <- (!train)
```

```{r}
regfit.best <- regsubsets(Salary ~., data = Hitters[train, ], nvmax = 19)
test.mat <- model.matrix(Salary ~., data = Hitters[test, ])
val.errors <- rep(NA, 19)
for(i in 1:19){
  coefi <- coef(regfit.best, id = i)
  pred <- test.mat[,names(coefi)]%*%coefi
  val.errors[i] <- mean((Hitters$Salary[test] - pred)^2)
}
val.errors
which.min(val.errors)
#The best model is the one that contains 10 variables
```
#We can create our own function to make this a lot easier!
```{r}
predict.regsubsets <- function(object, newdata, id, ...){
  form <- as.formula(object$call[[2]])
  mat <- model.matrix(form, newdata)
  coefi <- coef(object, id = id)
  xvars <- names(coefi)
  mat[,xvars]%*%coefi
}
```

```{r}
regfit.best <- regsubsets(Salary ~., data = Hitters, nvmax = 19)
coef(regfit.best, 10)
```
```{r}
k <- 10
set.seed(1)
folds <- sample(1:k, nrow(Hitters), replace = TRUE)
cv.errors <- matrix(NA, k, 19, dimnames = list(NULL, paste(1:19)))
for(j in 1:k){
  best.fit <- regsubsets(Salary ~., data = Hitters[folds != j,],nvmax = 19)
  for(i in 1:19){
    pred <- predict(best.fit, Hitters[folds == j, ], id = i)
    cv.errors[j,i] <- mean( (Hitters$Salary[folds == j] - pred)^2)
  }
}
```

```{r}
mean.cv.errors <- apply(cv.errors, 2, mean)
mean.cv.errors
which.min(mean.cv.errors)
#The minimum is 11.
```
#Cross validation selects an 11-variable model.
```{r}
#Perform best subset selection on the full data set in order to obtain the 11-variable model.
reg.best <- regsubsets(Salary ~., data = Hitters, nvmax = 19)
coef(reg.best, 11)
```

