
#Lending Club

#A.J. Imholte and Connor Doak

#What factors predict the risk of default? How can we create models to better loan to indivudals? What are some of the fundamentals of personal loans?

```{r}
load("Loans_100k.rda")
```
#Some Cool Graphs
```{r}
hist(Loans_100k$loan_amnt)
hist(Loans_100k$int_rate)
hist(Loans_100k$total_pymnt)
```

#Some More Cool Graphs
```{r}
boxplot(loan_amnt ~ grade, data = Loans_100k, main = "Loan Amount Based on Loan Grade")
boxplot(int_rate ~ grade, data = Loans_100k, main = "Loan Interest Rate Based on Loan Grade")
boxplot(total_pymnt ~ loan_status, data = Loans_100k, main = "Total Loan Payment given Loan Status")
boxplot(total_pymnt ~ grade, data = Loans_100k, main = "Total Loan Payment given Loan Grade")
boxplot(out_prncp ~ grade, data = Loans_100k, main = "Outstanding Principal given Loan Grade")
boxplot(out_prncp ~ loan_status, data = Loans_100k, main = "Outstanding Principal given Loan Status")
```
#Total Amount of Loans in Each Status Type
```{r}
table(Loans_100k$loan_status)
#as.integer(table(Loans_100k$loan_status))
```
#Median Interest Rates for Each Grade of Loan
```{r}
a_int_rate <- summary(Loans_100k$int_rate[Loans_100k$grade == "A"])
b_int_rate <- summary(Loans_100k$int_rate[Loans_100k$grade == "B"])
c_int_rate <- summary(Loans_100k$int_rate[Loans_100k$grade == "C"])
d_int_rate <- summary(Loans_100k$int_rate[Loans_100k$grade == "D"])
e_int_rate <- summary(Loans_100k$int_rate[Loans_100k$grade == "E"])
f_int_rate <- summary(Loans_100k$int_rate[Loans_100k$grade == "F"])
g_int_rate <- summary(Loans_100k$int_rate[Loans_100k$grade == "G"])

a_int_rate
b_int_rate
c_int_rate
d_int_rate
e_int_rate
f_int_rate
g_int_rate
```
Loan interest rates reflect risk.

#Exploratory regressions
```{r}
mod2 <- lm(int_rate ~ grade + loan_amnt + annual_inc + term, data = Loans_100k)
summary(mod2)
mod3 <- lm(total_pymnt ~ annual_inc + loan_amnt + grade + int_rate, data = Loans_100k)
summary(mod3)
mod4 <- lm(out_prncp ~ annual_inc + loan_amnt + grade + int_rate, data = Loans_100k)
summary(mod4)
```

```{r}
keepers <- Loans_100k$loan_status %in% c("Charged Off", "Fully Paid")
data_for_modeling <- Loans_100k[keepers,]
table(data_for_modeling$loan_status)
data_for_modeling$loan_status[data_for_modeling$loan_status == "Charged Off"] <- as.numeric(0)
data_for_modeling$loan_status[data_for_modeling$loan_status == "Fully Paid"] <- as.numeric(1)
#unique(data_for_modeling$loan_status)
```
#Our model on the whole dataset
```{r}
model_formula <- as.numeric(loan_status) ~ loan_amnt + int_rate + grade + annual_inc + grade + total_pymnt
glm_mod <- glm(model_formula, data = data_for_modeling, family = binomial)
summary(glm_mod)
```
```{r}
train <- 1:(nrow(data_for_modeling) / 2)
training_data <- data_for_modeling[train,]
testing_data <- data_for_modeling[-train,]
```
#Model on Training Data
```{r}
glm_mod_training <- glm(model_formula, data = training_data, family = binomial)
summary(glm_mod_training)
```
#Confusion Matrix
```{r}
probability <- predict(glm_mod_training, testing_data, type = "response")
table(testing_data$loan_status, probability > .95)
```
#Specificity, Sensitivity, and Test Error Rate
```{r}
test_error_rate <- (1792+70)/(2532+70+1792+9798)
specificity <- 2532/(2532+70)
#What percentage of the people that the model predicts would default that actually defaulted
sensitivity <- 9798/(1792+9798)
#What percentage of the people that would have fully paid that the model predicts would have paid
test_error_rate
specificity
sensitivity
```

#Loss Analysis
```{r}
payments <- median(data_for_modeling$total_rec_prncp)+median(data_for_modeling$total_rec_int)
amounts <- median(data_for_modeling$loan_amnt)
rate <- median(data_for_modeling$int_rate)
sd_payments <- sd(data_for_modeling$total_rec_prncp+data_for_modeling$total_rec_int)
sd_amounts <- sd(data_for_modeling$loan_amnt)
sd_payments
sd_payments/sd_amounts
#Variance is greater on payments than amounts
payments/amounts
#What fraction of loan amounts are paid
(payments * .94)*-.07
#The average loss of a loan, assuming that half of interest payments are paid and 95% of the principal
amounts*(1+rate/100) - amounts
#The average amount we will make per loan
```

```{r}
defaults <- (payments * .94)*-.07 * 70
#Expected value of defaults
expected_payment <- (amounts*(1+rate/100) - amounts)* (9798+70)
#The expected amount we will make in total if we loan to everyone
actual_payment <- expected_payment - defaults
actual_payment
defaults
```

#Another Confusion Matrix
```{r}
library(MASS)
lda_fit <- lda(model_formula, data = training_data)
lda_test <- predict(lda_fit, data = testing_data)$class
table(testing_data$loan_status, lda_test)
```

#Sensitivity, Specificity, and Test Error Rate
```{r}
test_error_rate <- (1376+2300)/(302+2300+1376+10214)
specificity <- 302/(302+2300)
sensitivity <- 10214/(10214+1376)
test_error_rate
specificity
sensitivity
```

#Risk Analysis
```{r}
defaults <- (payments * .94)*-.07 * 2300
#Expected value of defaults
expected_payment <- (amounts*(1+rate/100) - amounts)* (10214 + 2300)
#The expected amount we will make in total if we loan to everyone
actual_payment <- expected_payment - defaults
actual_payment
defaults
```

#QDA Confusion Matrix
```{r}
library(MASS)
qda_fit <- qda(model_formula, data = training_data)
qda_test <- predict(qda_fit, data = testing_data)$class
table(testing_data$loan_status, qda_test)
```

#Sensitivity, Specificity, and Test Error Rate
```{r}
test_error_rate <- (1987+2157)/(445+2157+1987+9603)
specificity <- 445/(2157+445)
sensitivity <- 9603/(1987+9603)
test_error_rate
specificity
sensitivity
```
#Risk Analysis
```{r}
defaults <- (payments * .94)*-.07 * 2157
#Expected value of defaults
expected_payment <- (amounts*(1+rate/100) - amounts)* (9603 + 2157)
#The expected amount we will make in total if we loan to everyone
actual_payment <- expected_payment - defaults
actual_payment
defaults
```
#Tree Classification Model
```{r}
library(tree)
loan_pure <- tree.control(nrow(training_data), mincut = 0, minsize = 1, mindev = 0)
status_classifier <- tree(model_formula, data = training_data, control = loan_pure)
status_classifier <- prune.tree(status_classifier, best = 15)
testing_status_classifier <- predict(status_classifier, testing_data)
testing_preds <- ifelse(testing_status_classifier > .80, 1, 0)
table(testing_data$loan_status, testing_preds)
```
#Test error rate, specificity, and sensitivity
```{r}
test_error_rate <- sum(71, 341) / sum(2261, 341, 71, 11519)
specificity <- 2261 / sum(2261, 341)
sensitivity <- 11519 / sum(71, 11519)
test_error_rate
specificity
sensitivity
```
#Risk analysis
```{r}
defaults <- (payments * .94)*-.07 * 2157
#Expected value of defaults
expected_payment <- (amounts*(1+rate/100) - amounts)* (9603 + 2157)
#The expected amount we will make in total if we loan to everyone
actual_payment <- expected_payment - defaults
actual_payment
defaults
```
