#A.J. Imholte
#K-fold Cross Validation on Logistic Regression
#Lending Club Project

```{r}
load("C:/Users/Aaron/Downloads/Loans.rda")
loan_sample <- data.frame(Loans)
```

```{r}
keepers <- loan_sample$loan_status %in% c("Charged Off", "Fully Paid")
data_for_modeling <- loan_sample[keepers, ]
data_for_modeling$loan_status[data_for_modeling$loan_status == "Charged Off"] <- as.numeric(0)
data_for_modeling$loan_status[data_for_modeling$loan_status == "Fully Paid"] <- as.numeric(1)
```

```{r}
data_for_modeling <- data_for_modeling[sample(nrow(data_for_modeling)),]
folds <- cut(seq(1,nrow(data_for_modeling)),breaks=10,labels=FALSE)
formula3 <- as.numeric(loan_status) ~ loan_amnt + int_rate + grade + annual_inc + total_pymnt + sub_grade + funded_amnt_inv + verification_status + dti + open_acc + revol_bal
```

#10-k Fold Cross Validation on Logistic Regression Model
```{r}
set.seed(10)
for(i in 1:10){
    testIndexes <- which(folds==i,arr.ind=TRUE)
    testData <- data_for_modeling[testIndexes, ]
    trainData <- data_for_modeling[-testIndexes, ]
    glm_fit <- glm(formula3,data = trainData, family = binomial)
    glm_test <- predict(glm_fit, testData, type = "response")
    print(table(testData$loan_status, glm_test > .7))
}
```

```{r}
spec1 <- 4287/sum(4287,282)
spec2 <- 4267/sum(4267,313)
spec3 <- 4197/sum(4197,308)
spec4 <- 4196/sum(4196,362)
spec5 <- 4124/sum(4124,306)
spec6 <- 4207/sum(4207,319)
spec7 <- 4191/sum(4191,306)
spec8 <- 4104/sum(4104,336)
spec9 <- 4233/sum(4233,348)
spec10 <- 4235/sum(4235,327)
spec.cv <- mean(spec1, spec2, spec3, spec4, spec5, spec6, spec7, spec8, spec8, 
                spec10)
```

#10-k Fold Cross-Validated Specificity
```{r}
spec.cv
```

```{r}
sens1 <- 20657/sum(72,20657)
sens2 <- 20650/sum(67,20650)
sens3 <- 20708/sum(84,20708)
sens4 <- 20716/sum(94,20716)
sens5 <- 20805/sum(62,20805)
sens6 <- 20707/sum(64,20707)
sens7 <- 20725/sum(75,20725)
sens8 <- 20772/sum(85,20772)
sens9 <- 20631/sum(85,20631)
sens10 <- 20656/sum(79,20656)
sens.cv <- mean(sens1, sens2, sens3, sens4, sens5, sens6, sens7, sens8, sens9, sens10)
```

#10-k Fold Cross-Validated Sensitivity
```{r}
sens.cv
```

```{r}
classErr1 <- sum(72, 282)/25298
classErr2 <- sum(67, 313)/25297
classErr3 <- sum(84, 308)/25297
classErr4 <- sum(67, 362)/25297
classErr5 <- sum(62, 306)/25297
classErr6 <- sum(64, 319)/25297
classErr7 <- sum(75, 306)/25297
classErr8 <- sum(85, 336)/25297
classErr9 <- sum(85, 348)/25297
classErr10 <- sum(79, 327)/25297
classErr.cv <- mean(classErr1, classErr2, classErr3, classErr4, classErr5, classErr6, classErr7, classErr8, classErr9, classErr10)
```

#10-k Fold Cross-Validated Classification Error
```{r}
classErr.cv
```

K-fold cross validation (k = 10) results in a specificity of 94%, a sensitivity of 99.6%, and an overall classification error of 1.4% for the proposed logistic regression model.
